---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

1. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
totalSteps<-aggregate(steps ~ date,data = activity,sum,na.rm = TRUE)
```

## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day
```{r}
hist(totalSteps$steps)
```

2. Calculate and report the **mean** and **median** total number of
   steps taken per day

```{r echo=TRUE}
mean(totalSteps$steps)
median(totalSteps$steps)
```

## What is the average daily activity pattern?

* Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
stepsInterval<-aggregate(steps ~ interval,data = activity,mean,na.rm = TRUE)
plot(steps ~ interval,data = stepsInterval,type = "l")
```

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps? 
```{r}
stepsInterval[which.max(stepsInterval$steps),]$interval
```

It is the **`r stepsInterval[which.max(stepsInterval$steps),]$interval`th** interval.


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
